---
title: "Chena Drift Permit Reporting"
output:
  html_document: 
    df_print: paged
    fig_width: 10
    fig_height: 6
    fig_caption: yes
    code_folding: hide
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: inline
---
Website last updated `r Sys.time()` by Erik Schoen (eschoen@alaska.edu)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r initialize script, include=FALSE}

# clear environment
rm(list=ls())

#require packages
library(googlesheets)
library(tidyverse)
library(hms)
library(janitor)
library(lubridate)
library(readxl)
library(DT)
library(readr)

# set plotting theme
theme_set(theme_bw(12))

# if editing this document, set user to access source files
#user <- "/Users/bmeyer/"

# set data source directory for fish data
#source_directory <- paste0(user,"Dropbox/Chena_Data_2020/Chena Drift Project Data_SFR/UAF Chena Drift Data/Fish/")

```

<br>

### Introduction

This document will generate permit reports from the Chena Drift/Fire project 2019-2020

* ADFG
* INAD
* IACUC

<br>

### Import data

Click on "Code" to see import process
```{r}
# read in data from dropbox directory

## 2019
sites2019 <- read_excel("data/Fish/2019/2019 Data Inputs/Chinook catch 2019.xlsx", sheet = "Sites") %>%
  mutate(`General Area` = "Upper Chena Mainstem")
fish2019 <- read_excel("data/Fish/2019/2019 Data Inputs/Chinook catch 2019.xlsx", sheet = "Fish")

## 2020
sites2020 <- read_excel("data/Fish/2020/2020 Data Inputs/Fishing Data 2020.xlsx", sheet = "Fishing_Sites") %>%
  select(GPSwpt, Lat, Lon, `General Area`)
fish2020 <- read_excel("data/Fish/2020/2020 Data Inputs/Fish Datasheet 2020.xlsx", skip = 3)

# Clean up and summarize 2019 data----------
## Convert 2019 data date to date format and join with site data
fish2019  <- fish2019 %>%
  mutate(Date = ymd(Date)) %>%
  left_join(sites2019, by = "GPSwpt") %>%
  select(-DatasheetOrder,-TL_mm,-Description) %>%
  rename(General.Area = `General Area`)

# General Note: do not alter 2019 data formats, because J. Neuswager and E. Schoen may be already using these sheets as is for other script analyses.  Instead, make 2020 data conform to 2019 formats.

# reformat 2020 data to join with 2019 data
fish2020 <- fish2020 %>%
  rename(GPSwpt = Site,
         ObserverName = Observer,
         FL_mm = `FL (mm)`,
         Wt_g = `Wt (g)`,
         Mortality = `Mort?`) %>%
  select(-`Disposition (MR/ID)`, -`Data Entry`,-`Addl. Count`,-Count) %>%
  left_join(sites2020, by = "GPSwpt") %>%
  transform(FL_mm = as.numeric(FL_mm),
            Wt_g = as.numeric(Wt_g),
            Lat = as.character(Lat),
            Lon = as.character(Lon))  

# join 2019 and 2020 data
fish <- bind_rows(fish2019,fish2020) %>%
  mutate(year = year(Date)) 
```

<br>

#### Overall catch by date/species
```{r}
datatable(totalCatchByDate <- fish %>%
  group_by(Species, Date) %>%
  summarize(TotalCatch = sum(Number)) %>%
  arrange(Species,Date))
```

<br>

#### Overcall catch by species and year
```{r}
# Check how many of each species were captured
datatable(totalCatch <- fish %>%
  group_by(Species,year) %>%
  summarize(TotalCatch = sum(Number)) %>%
  arrange(year,Species))

```

***

### INAD reporting 2020

```{r}
# INAD-----(did this in Excel in 2019, but done this in R for 2020. B Meyer copied example from 2019 Excel format, found at ./Fish/Permit Reports/2019 Permit Reports) to generate R results here.

# 2020 INAD report
# create general summary table
inad.table <- fish %>%
  filter(year == 2020) %>%
  group_by(Date,GPSwpt,Species,Number,`Aqui.S.bath.`) %>%
  summarise(Fish.Per.Treatment = sum(Number)) %>%
  ungroup() %>%
  select(-Number)

# a fresh aqui-s bath was mixed at each unique instance of site/date where Chinook and/or chum were
# captured.  Create table describing creation order of aqui-s solution batches.
treatments <- fish %>%
  filter(year == 2020,
         Species %in% c("Chinook","Chum")) %>%
  group_by(Date,GPSwpt) %>%
  summarise(Fish.Per.Treatment = sum(Number)) %>%
  ungroup() %>%
  mutate(Treatment.Number = row_number()) %>% 
  select(-Fish.Per.Treatment) 
```

<br>

#### INAD Summary Table 1

Aqui-S use by Species, Site, and Date
```{r}
# join tables and create new columns to create INAD Summary Table 1
## grouped by species, site, and date
datatable(inad.summary.table.1 <- left_join(inad.table,treatments) %>%
            #filter(Species == "Chinook") %>%
  mutate(`Level of Anesthesia` = "Light (measure and weigh)",
         `Dose (mg/L)` = 20,
         `Time To Anesthesia (Min:Sec)` = "3:00",
         `Time To Recover (Min)` = "5:00",
         Principal_Investigator = "Erik Schoen (eschoen@alaska.edu")) 

# save summary table 1
write.csv(inad.summary.table.1,"output/tables/permit reports/AquiS_BySite_2020.csv",row.names = F)

```

<br>

#### INAD Summary Table 2

Aqui-S use by Site and Date
```{r}
# create INAD Summary Table 2
## calculate daily quantities of aqui-s used at X number of sites
datatable(inad.summary.table.2 <- inad.summary.table.1 %>%
  distinct(Date,GPSwpt) %>%
  group_by(Date) %>%
  summarise(number_sites = n()) %>%
  mutate(`Total Quantity Aqui-S Used (mL)` = number_sites,
         Principal_Investigator = "Erik Schoen (eschoen@alaska.edu"))

# save
write.csv(inad.summary.table.2,"output/tables/permit reports/AquiS_BySite_2020.csv",row.names = F)

```

<br>

```{r include=F}
(quant2020 <- sum(inad.summary.table.2$`Total Quantity Aqui-S Used (mL)`))
```

* **Total quantity Aqui-S used in 2020 was `r quant2020` mL.**

<br>

***

### ADFG Fish Use Permit Reports data

#### Daily and annual mortality 

##### 2019

```{r}

# ADFG-------
# Create report for Fish Use Permit
# Check whether an instance of > 10% unintended collecting mortality occurred (requires contacting AMB)
datatable(mort19 <- fish %>%
  filter(year == 2019) %>%
  ungroup() %>%
  group_by(Date, Mortality,year) %>%
  summarize(Total = sum(Number)) %>%
  spread(Mortality, Total) %>%
  rename(N = `<NA>`) %>%
  mutate(pct_mort = (Y / N)*100))

# summarise overall mortality rate
(mort19_overall <- mort19 %>%
  ungroup() %>%
  summarise(sum_mort = sum(Y, na.rm = T),
            sum_non_mort = sum(N)) %>%
  mutate(overall_mort_pct = (sum_mort / sum_non_mort)*100))
```

<br>

* **For 2019, the max daily mortality rate was 2.72% (7/26/19), and overall 2019 rate was 1.6%**

<br>

##### 2020

```{r}
# calculate 2020 mortality rates
datatable(mort20 <- fish %>%
    filter(year == 2020) %>%
    ungroup() %>%
    group_by(Date, Mortality,year) %>%
    summarize(Total = sum(Number)) %>%
    spread(Mortality, Total) %>%
    rename(N = `<NA>`) %>%
    mutate(pct_mort = (Y / N)*100))

# summarise overall mortality rate
(mort20_overall <- mort20 %>%
    ungroup() %>%
    summarise(sum_mort = sum(Y, na.rm = T),
              sum_non_mort = sum(N)) %>%
    mutate(overall_mort_pct = (sum_mort / sum_non_mort)*100))

```

<br>

* **For 2020, the max daily mortality rate was 1.67% (6/3/20), and overall 2020 rate was 0.15%**

<br>

***

#### Aquatic Resource Permit reports 

##### 2019
```{r}
# Generate output table for 2019 ADFG ARP report
datatable(adfg19 <- fish %>%
  filter(year == 2019) %>%
  mutate(Location.ID = NA,
         Latitude = Lat,
         Longitude = Lon,
         Datum = "WGS84",
         Coordinate.determination.method = "MAP",
         Name.of.water.body = "Chena River",
         Date = format(Date, "%m/%d/%Y"),
         Observer.name = ObserverName,
         Fish.collection.method = Gear,
         Species = ifelse(Species == "Chinook", "Chinook salmon",
                          ifelse(Species == "Chum", "chum salmon",
                                 ifelse(Species == "Sculpin", "Slimy sculpin", Species))),
         Life.stage = "juvenile",
         Length.mm = FL_mm,
         Length.method = ifelse(is.na(Length.mm), NA, "Fork"),
         Weight.g = Wt_g,
         Sex = NA,
         Age = NA,
         Age.method = NA,
         ADFG.GCL = NA,
         Additional.count.1 = ifelse(Number == 1, NA, Number),
         Disposition.1 = ifelse(is.na(Mortality),
                                ifelse(is.na(Length.mm), "ID'ed and Released", 
                                       "Measured and Released"),
                                "Unintended Mortality"),
         Additional.count.2 = NA,
         Disposition.2 = NA,
         Anesthesia_Sedative.Used = ifelse(is.na(Length.mm), NA, "Aqui-S 20E"),
         Comments = NA) %>%
  select(Location.ID:Name.of.water.body, Date, Observer.name, Fish.collection.method, Species, 
         Life.stage:Comments))

# write 2019 ADFG report data to local dropbox directory
write_excel_csv(adfg19, "output/tables/permit reports/ADFG Permit Data 2019.csv", na = "")

```

<br>

##### 2020
```{r}

# Generate output table for 2020 ADFG ARP report
datatable(adfg20 <- fish %>%
  filter(year == 2020) %>%
  mutate(Location.ID = NA,
         Latitude = Lat,
         Longitude = Lon,
         Datum = "WGS84",
         Coordinate.determination.method = "MAP",
         Name.of.water.body = "Chena River",
         Date = format(Date, "%m/%d/%Y"),
         Observer.name = ObserverName,
         Fish.collection.method = Gear,
         Species = ifelse(Species == "Chinook", "Chinook salmon",
                          ifelse(Species == "Chum", "chum salmon",
                                 ifelse(Species == "Sculpin", "Slimy sculpin", Species))),
         Life.stage = "juvenile",
         Length.mm = FL_mm,
         Length.method = ifelse(is.na(Length.mm), NA, "Fork"),
         Weight.g = Wt_g,
         Sex = NA,
         Age = NA,
         Age.method = NA,
         ADFG.GCL = NA,
         Additional.count.1 = ifelse(Number == 1, NA, Number),
         Disposition.1 = ifelse(is.na(Mortality),
                                ifelse(is.na(Length.mm), "ID'ed and Released", 
                                       "Measured and Released"),
                                "Unintended Mortality"),
         Additional.count.2 = NA,
         Disposition.2 = NA,
         Anesthesia_Sedative.Used = ifelse(is.na(Length.mm), NA, "Aqui-S 20E"),
         Comments = NA) %>%
  mutate(Observer.name = str_replace(Observer.name, "ES","Erik Schoen"),
         Observer.name = str_replace(Observer.name, "JH","Justin Hill"),
         Observer.name = str_replace(Observer.name, "MQ","Michelle Quillen"),
         Observer.name = str_replace(Observer.name, "ED","Erick De La Rosa"),
         Observer.name = str_replace(Observer.name, "EDLR","Erick De La Rosa"),
         Observer.name = str_replace(Observer.name, "RosaLR","Rosa"),
         Observer.name = str_replace(Observer.name, "WS","Will Samuel"),
         Observer.name = str_replace(Observer.name, "BM","Benjamin Meyer"),
         Fish.collection.method = str_replace(Fish.collection.method, "MT","Minnow Trap")) %>%
  select(Location.ID:Name.of.water.body, Date, Observer.name, Fish.collection.method, Species, 
         Life.stage:Comments))

# Note : coordinates are not included for South Fork MT 1 and South Fork MT 2 sites.  Unable to diagnose as of 9/25/20.  Fix manually in excel spreadsheet for now. Excel spreadsheet created on 9/25/20 is up to date.

# write 2020 ADFG report data to local dropbox directory
# write_excel_csv(adfg20, "Permit Reports/2020 Permit Reports/ADFG Permit Data 2020.csv", na = "")
```

<br>

***

### IACUC

Generate summary data for IACUC report
```{r}
# IACUC--------
# Calculate how many Chinook were "used", i.e., anesthetized, measured, and released. Don't count
# the fish that were only ID'd and released (but do count all unintended mortalities even if not
# measured).
datatable(iacuc <- bind_rows(adfg19,adfg20) %>%
  transform(Date = mdy(Date)) %>%
  mutate(year = year(Date),
         protocol = ifelse(Date >= ymd(20200826), "Chena Fire", "Chena Drift")) %>%
  filter(Species == "Chinook salmon" | Species == "Slimy sculpin") %>%
  # Remove the fish that were not measured or weighed or unintended morts
  # filter(!is.na(FL_mm) | !is.na(Wt_g) | Mortality == "Y") %>% 
  mutate(Number = ifelse(is.na(Additional.count.1), 1, Additional.count.1)) %>%
  group_by(Disposition.1,year, protocol) %>%
  summarize(Count = sum(Number, na.rm = T)))

#save
write_csv(iacuc,"output/tables/permit reports/iacuc_permit_data.csv")

# Calculte mortality rates by IACUC protocol (Chinook only)
# Chena drift 2020:
mort.drift.2020 <- 1/634
# Chena drift project had < 0.2% mortality (1/634)
# Chena fire project had zero mortality
```


