---
title: "Chena Fire & Drift\n Exploratory Data Analysis 2020"
output:
  html_document: 
    df_print: paged
    fig_width: 10
    fig_height: 6
    fig_caption: yes
    code_folding: hide
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: false
editor_options: 
  chunk_output_type: inline
---
Website last updated `r Sys.time()` by Benjamin Meyer (bemeyer@alaska.edu)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r initialize script, include=FALSE}

# clear environment
rm(list=ls())

#require packages
library(googlesheets)
library(tidyverse)
library(hms)
library(janitor)
library(lubridate)

```

<br>

<iframe src="https://drive.google.com/file/d/1xyt8j4ivxhsnB0CxEd2rRo2Zttp8Gqjl/preview" width="640" height="480"></iframe>

*Headwater tributary of the Chena River in the 2019 Nugget Creek Fire scar.  Photo from July 2020.*

***

<br>

### Introduction

This document will outline preliminary results from exploratory data analysis from UAF/SFR's ongoing Chena Fire/Drift 2020 research

***

<br>

### People

##### Principal Investigator

* Dr. Erik Schoen, UAF (https://www.researchgate.net/profile/Erik_Schoen)

##### Co-Principal Investigators

* Dr. Jason Neuswanger, SFR (http://www.jasonneuswanger.com/)
* Dr. Carol Volk, SFR (https://southforkresearch.org/)
* Dr. Mark Wipfli, UAF (http://people.iab.uaf.edu/mark_wipfli)
* Brian McKenna, TCC

##### Other Personnel

* Benjamin Meyer, UAF (https://www.researchgate.net/profile/Benjamin_Meyer8)
* Erick De La Rosa, UAF
* Michelle Quillen, UAF
* Will Samuel, UAF
* Justin Hill, SFR
* Debra Lynne, TCC

##### Useful Links

* Overview of precision drift sampling fieldwork

  + http://www.southforkresearch.org/2020_field_tech.html
  + https://www.youtube.com/watch?v=5whZhZUvwdo&feature=youtu.be

<br>

***

<br>

### Methods

##### Study Area

* Map of study area and field sites

  + View map on ArcGIS Online at https://arcg.is/04brTu
  + View coordinates for all sites at http://bit.ly/chenadriftcoordinates

<style>.embed-container {position: relative; padding-bottom: 80%; height: 0; max-width: 100%;} .embed-container iframe, .embed-container object, .embed-container iframe{position: absolute; top: 0; left: 0; width: 100%; height: 100%;} small{position: absolute; z-index: 40; bottom: 0; margin-bottom: -15px;}</style><div class="embed-container"><iframe width="400" height="300" frameborder="0" scrolling="no" marginheight="0" marginwidth="0" title="Chena Drift/Fire" src="//www.arcgis.com/apps/Embed/index.html?webmap=5f7eef523b844314b79085429a72cd89&extent=-147.5065,64.6816,-146.2033,64.9876&zoom=true&previewImage=false&scale=true&disable_scroll=true&theme=light"></iframe></div>

***
    
<br>

##### Field Sampling Effort

<br>

When and where did in-situ aquatic ecology data collection occur, and by whom?

  + Main stem drift (SFR 2019-2020, UAF 2020)
```{r}
# Will import from J. Neuswanger Python output results.  B Meyer will use to create/display summary table of sampling efforts here
```
  + Headwaters drift (UAF 2020)
```{r}
# Will import from J. Neuswanger Python output results.  B Meyer will use to create/display summary table of sampling efforts here
```
   + Juvenile Chinook fishing effort (SFR 2019-2020, UAF 2020)
```{r}
# Import Excel sheet from Dropbox, "Fishing Data 2020.xlsx".  
# a.) Create simple bar chart of catch by site faceted by week
# b.) Create plot of CPUE vs time for MT and dipnets, faceted by Upper vs. Lower river
```
  + Water temperature logger data, (UAF/SFR 2019 & 2020)
```{r}
# Import Excel sheet from dropbox, "/Users/bmeyer/Dropbox_auxiliary/Dropbox/Chena_Data_2020/UAF_Data/2020_Logger_Downloads/HOBO_Final_Data".  Create summary figure displaying temporal extent of various loggers and their parameters.
```
  + Hydrolab Sonde turbidity, DO, conductivity, and water temperature data, (UAF 2008, 2009, 2020)
```{r}
# Import Excel sheet from dropbox, "/Users/bmeyer/Dropbox_auxiliary/Dropbox/Chena_Data_2020/UAF_Data/2020_Logger_Downloads/Hydrolab_Final_Data"
# read in 2020 Hydrolab data in progress

```
  
<br>  

##### Remote Sensing Data collection

When and where did remote sensing data collection occur, and by whom?  
  
  + Drone aerial imagery (UAF/TCC 2020)
    + Drone Footage from 6/29/2020 of Chena Mainstem Drift Sampling [https://drive.google.com/drive/u/0/folders/1xaSedWQSAMwjPAbvkAYaH8gsdCRZUjoZ]
    
```{r}
# Summary table of dates & extent of aerial imagery data collected
```
  + Remote sensing of watershed vegetation composition and % of watershed burned (UAF 2020)
```{r}
# Currently in discussion with potential collaborator(s) Dr. Alison York and Dr. Santosh Pandya
```






<br>
  
##### Laboratory Data Collection Effort

How many samples were processed, and what type?

  + Coarse fraction
```{r}
# share in-text result from J. Neuswangers Python output table
```
  + Fine Fraction
```{r}
# share in-text result from J. Neuswangers Python output table
```

<br>

***

<br>

### Results

##### Headwaters Tributaries

* Were habitat conditions distinct in recent burn vs. old burn tributaries?

  + Parameters
    + Drift debris density
    + Invertebrate density
    + Prey energy density
    + Water temperature
    + Turbidity
    + Others
    
    
##### Chena Mainstem

* Were habitat conditions distinct in the main stem upstream vs. downstream of the 2019 Nugget Creek fire?

  + Parameters
    + Drift debris density
    + Invertebrate density
    + Prey energy density
    + Water temperature
      + Preliminary draft results at https://rpubs.com/bmeyer/watertemps1
    + Turbidity
      + Preliminary draft results at https://rpubs.com/bmeyer/sondes2020
    + Juvenile Chinook salmon growth rate
      + Preliminary draft at https://rpubs.com/bmeyer/chena_chinookgrowth
```{r}
# Import  Excel sheet from dropbox, "Fish Datasheet 2020.xlsx".  Create plot(s) of mean FL vs time, a.) all sites combined and b.) by lower vs. upper river c.) by year
```
  
***

### Photos

Various fieldwork photos found at: https://drive.google.com/drive/folders/11uDSnF9GssjUBe2RUwTEeC1RAgKMVY2p?usp=sharing 

<br>

### Permit Reports Data

Find more at https://rpubs.com/bmeyer/chenafiredriftpermits
